---
title: "Regression on flight data"
output: html_document
---

Get access of the data /
https://drive.google.com/file/d/1-bBXcpZZ0XbRui4825J5SVghs3Zhu16T/view?usp=sharing \
Data includes 2019-2021 US airlines including American Airlines, United Airlines and Delta Airlines\ 

The original data was provided by ICAO: \
https://www.transtats.bts.gov/Tables.asp?QO_VQ=EFD&QO_anzr=Nv4yv0r%FDb0-gvzr%FDcr4s14zn0pr%FDQn6n&QO_fu146_anzr=b0-gvzr \


```{r setup}
library(tidyverse)
library(here)
library(readr)
library(future.apply)
library(dplyr)
library(httr)
library(jsonlite)
library(RColorBrewer)
library(scales)
library(feather)
library(caret)

```

read file 
```{r}
raw_data <- read_feather(here('mlm_dataset_5.feather'))
```

Correlation plot 
```{r}
library(corrplot)
num_col <- select_if(raw_data, is.numeric)

# Dataset too large, thus need sampling 
num_col <- num_col %>% sample_n(50000)

corrplot(cor(num_col), method = 'number')
corrplot(cor(num_col), method = 'ellipse')

```

Chi-square independence test 
```{r}
library(broom)
cat_col <- raw_data %>% select_if(negate(is.numeric))

#First column against other columns 
CHIS <- lapply(cat_col[,-1], function(x) chisq.test(cat_col[,1], x))
chi_result <- do.call(rbind, lapply(CHIS,tidy))

show(chi_result)
```

Point-biseral test 

When cor() function in R applied to a numerical and a dichotomous variable, it's the equivalence of point-biseral test
```{r}
#Turn cateogrical variables into dichotomous

#requires a lot of memory to run dummy variable codes 
small_data <- raw_data %>% sample_n(50000)
  
dmy <- dummyVars(" ~ . ", data= small_data, fullRank = T)
point_test_data <- data.frame(predict(dmy, newdata = small_data))

#correlation test (including everything)
corr_result_all <- round(cor(point_test_data), digits = 2)

show(corr_result_all)
```
For interpreting the result, look at the results among numerical and categorical variables (after dummy). 


Numerical data transformation
Min-max normalization 
```{r}
min_max_norm <- function(x){
  (x-min(x)) / (max(x) - min(x))
}

num_col <- select_if(raw_data, is.numeric)

num_col_norm <- as.data.frame(lapply(num_col, min_max_norm))
```

Regrouping data
```{r}
cat_col <- raw_data %>% select_if(negate(is.numeric))
num_col_norm$id <- seq.int(nrow(raw_data))
cat_col$id <- seq.int(nrow(raw_data))
combine_data <- left_join(num_col_norm, cat_col, by = 'id')
normalized_data <- combine_data %>% select(-id)
mlm_all<- normalized_data
```

linear regression 
```{r}
mlm1 <- lm(dep_delay ~ month + day_of_week + dest_state_abr + dep_time +
             airline + taxi_out + origin_airport_code + dest_airport_code +
             air_time + taxi_in + manufacturer + model + age +
             emptotal, data = mlm_all)
summary(mlm1) # R-Squared 0.03944 p < 2.2e-16
plot(mlm1)
```

Boruta feature selection
```{r}
# Run Boruta
boruta_sasuke <- Boruta(dep_delay ~ month + day_of_week + dep_time +
                        airline + taxi_out + origin_airport_code + 
                        model + age, data = mlm_all, doTrace = 2)

boruta_signif <- names(boruta_sasuke$finalDecision[boruta_sasuke$finalDecision %in% c("Confirmed", "Tentative")])
print(boruta_signif)
plot(boruta_sasuke, cex.axis = .7,
     las = 2, xlab ="", main = "Variable Importance")

```

New linear regression model based on Boruta feature selection
```{r}
mlm5 <- lm(dep_delay ~ month + dep_time +
             airline + age, data = mlm_all)
summary(mlm5) # R-Squared .01202 p-value <2.2e-16
VIF(mlm5)
```

Add internaction variables 
```{r}
mlm_7 <- lm(dep_delay ~ month + dep_time +
              airline + age + month:age + 
              airline:age + airline:dep_time
            , data = mlm_all)

summary(mlm_7)
```
Interaction variable not significant --> no need to add interaction variable

Final model is mlm5: diagnostic plots 
```{r}
# Model diagnoistic 

pairs(mlm_CA)
corrplot(cor(mlm_CA_sub3), method = "number")


# Residual Plot on MLM 5

plot(mlm5)
```

Binomial regression
1. filter data those with positive depature delays
```{r}
blm_bino <- mlm_all %>% 
  filter(dep_delay > 0)
summary(blm_bino)
#dep delay median is 12, mean 20.8. 1Q 4, 3Q 30
```

2. Create new Binomial delay 
```{r}
blm_bino <- blm_bino %>% 
  mutate(delay_new = dep_delay > 12)

summary(blm_bino)
#344677 true, #356513 false
```

3. Train/Test set data
```{r}
blm_bino$id <- 1:nrow(blm_bino)
blm_bino.train <- blm_bino %>% dplyr::sample_frac(0.75)
blm_bino.test <- dplyr::anti_join(blm_bino, blm_bino.train, by ='id')
```

4. GLM model 
```{r}
glm0 <- glm(delay_new ~ month + day_of_week + dest_state_abr + dep_time +
              airline + taxi_out + origin_airport_code + dest_airport_code +
              air_time + taxi_in + manufacturer + model + age +
              emptotal, family = binomial(logit), data = mlm_bino.train)
summary(glm0)

glm1 <- glm(delay_new ~ month + dep_time +
              airline + taxi_out + 
              model + age, family = binomial(logit), data = mlm_bino.train)
summary(glm1)
```

5. Model only for American Airlines
```{r}
blm_aa.train <- blm_bino.train %>% 
  filter(airline == 'American Airlines Inc.')

blm_aa.test <- blm_bino.test %>% 
  filter(airline == 'American Airlines Inc.')

blm2 <- glm(delay_new ~ month + dep_time +
              model, 
              family = binomial(logit), data = blm_aa.train)
summary(blm2)
```

6. Model only for Delta 
```{r}
blm_dl.train <- blm_bino.train %>% 
  filter(airline == 'Delta Air Lines Inc.')

blm_dl.test <- blm_bino.test %>% 
  filter(airline == 'Delta Air Line Inc.')

blm_ua.train <- blm_bino.train %>% 
  filter(airline == 'United Air Lines Inc.')

blm_ua.test <- blm_bino.test %>% 
  filter(airline == 'United Air Line Inc.')


blm3 <- glm(delay_new ~ month + dep_time +
              + model, 
            family = binomial(logit), data = blm_dl.train)
summary(blm3)

blm4 <- glm(delay_new ~ month + dep_time +
              model, 
            family = binomial(logit), data = blm_ua.train)
summary(blm4)
```

VIF of models 
```{r}
VIF(glm1)
VIF(blm2)
VIF(blm3)
VIF(blm4)

1/VIF(blm2)
```

Model validation
```{r}
probability_aa <- predict(blm2, newdata = blm_aa.test, type='response')
prediction_aa <- ifelse(probability_aa > 0.5, 1, 0)
confusion_aa <- table(blm_aa.test$delay_new, prediction_aa)

confusion_aa
```

Confusion Matrix 
```{r}
library(DHARMa)

res <- simulateResiduals(blm2, plot = T)

# Table

library(knitr)
kable(confusion_aa, caption = "Confusion Matrix")
```

